Disturbance-aware minimum-time planning strategies for motorsport vehicles with probabilistic safety certificates

### 2025-06-03 performance_comparison.tex
[MM] alignment
[mg] subsec 5.1

The second parameter examined is the initial value of the covariance matrix, $\bP_0$. From this matrix, the standard deviation of each component of the mean state vector can be computed as:
\begin{equation}
	\sigma_i = \sqrt{e_i^T \bP_0 e_i},\label{eq:sigma_from_P}
\end{equation}
where $e_i$ is the unit vector with a 1 in the $i$-th position and zeros elsewhere. These standard deviations, $\sigma_i$, are used to evaluate the probability of the state vector remaining within a given interval. For example, considering the first element of the state vector, $\mu_1$, the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ corresponds to a 68\% probability of containing the true state, while the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ corresponds to approximately 99\%.

In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$. These correspond, respectively, to doubling and halving the initial uncertainty associated with the state vector, as it's possible to verify from \eqref{eq:sigma_from_P}. 
For simplicity, a diagonal $\bar{\bP}_0$ has been chosen for this study. In this case, $\bar{\bP}_0$ is constructed by assigning the standard deviations of the elements of the state vector---collected in the vector $\bar{\boldsymbol{\sigma}}_0$---and placing their squared values along the diagonal. 
The values of $\bar{\boldsymbol{\sigma}}_0$ has been chosen accordingly to the range of variation of the associated variable. In particular, assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, we selected $\bar{\boldsymbol{\sigma}}_0 = \left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.

The results of the analysis conducted on this parameter are shown in the right panel of Figure~\ref{fig:ol_sensitivities} where the black line corresponds to the baseline, obtained setting $\bP_0 = \bar{\bP}_0$, while the red and the blue lines corresponds to configurations with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively. 
Obviously, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term. This is reflected in the blue line, which travels closer to the centerline w.r.t. the other two lines.

@@

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented along the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.

>>>>

The second parameter examined is the initial value of the covariance matrix, $\bP_0$. 
We recall that at each grid point a new prediction horizon is initialized, and the corresponding matrix $\bP$ at its starting point, i.e., $\bP^0_k$, is set to $\bP_0$. 
The diagonal elements of $\bP$ are the variances $\text{var}(x_i)$ of each state variables, while the off-diagonal elements encode their covariances $\text{cov}(x_i,x_j)$. Larger diagonal values indicate higher uncertainty in the corresponding states, and nonzero off-diagonal terms imply mutual dependence between them. For simplicity, a diagonal $\bP_0$ is used in this study, that is, for each prediction horizon we assign an initial standard deviation $\sigma_i = \sqrt{\text{var}(x_i)}$ to each state and assume all initial covariances to be zero.

To provide a concrete interpretation, consider the first state variable $x_1$ with mean $\mu_1$ and standard deviation $\sigma_1$.
Under the Gaussian assumption, the true value of $x_1$ lies within the interval $\left[\mu_1 - \sigma_1,\ \mu_1 + \sigma_1\right]$ with approximately 68\% probability, and within the interval $\left[\mu_1 - 3\sigma_1,\ \mu_1 + 3\sigma_1\right]$ with approximately 99\% probability.

We assign a baseline initial covariance matrix $\bP_0 = \bar{\bP}_0$ by selecting standard deviations $\bar{\sigma}_i$ for each state variable. These standard deviations have been chosen based on the typical range of variation of each state.
Assuming the state vector is ordered as described in Section~\ref{sec:vehicle_model}, the values used are $\bar{\boldsymbol{\sigma}}~=~\left[0.1\,\textrm{m/s}, 0.01\,\textrm{m/s}, 0.01\,\textrm{rad/s}, 1\,\textrm{m}, 1\,\textrm{m}, 0.0175\,\textrm{rad}\right]^T$.

In addition to the baseline value $\bar{\bP}_0$, two alternative values are considered: $\bP_0 = 4\bar{\bP}_0$ and $\bP_0 = \frac{1}{4}\bar{\bP}_0$. These correspond, respectively, to doubling and halving the initial standard deviations $\sigma_i$ associated with the state vector. 
A comparison between the trajectories obtained using different values of $\bP_0$ is shown in the right panel of Figure~\ref{fig:ol_sensitivities}.
The black line corresponds to the baseline case with $\bP_0 = \bar{\bP}_0$, while the red and the blue lines represent the cases with $\bP_0 = \frac{1}{4}\bar{\bP}_0$ and $\bP_0 = 4\bar{\bP}_0$, respectively. As reasonably expected, using a larger initial covariance matrix results in a greater covariance matrix after $H$ steps, which in turn leads to a higher back-off term. This is reflected in the blue line, which travels closer to the centerline w.r.t. the other two lines.

@@

\subsection{Simulations with noise realization}
This analysis aims to provide empirical validation of the robust trajectories obtained using the methods proposed in this work. For conciseness, the comparison is limited to a nominal trajectory and a trajectory computed using the closed-loop method with only the track limit constraint robustified.
An LQR controller, designed according to the procedure outlined in Section~\ref{sec:LQR}, is implemented to stabilize the nominal trajectory. In contrast, the robust trajectory is tracked using the controller directly obtained from the optimization process.
